---
title: "STAT545B - AssignmentB-1"
output: github_document
author: "Yue Shen"
date: "2021/11/02"
---


# Welcome to Yue Shen's assignmentB-1!

Loading data, tidyverse, and testthat package below: 

```{r, message = FALSE}
suppressPackageStartupMessages(library(datateachr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(testthat))
```


## Exercise 1: Make a Function (25 points)

## Exercise 2: Document your Function (20 points)

When looking back to the previous mini-data-analysis-project, I found out that I usually needed to compute some summary statistics(such as max, min and mean) of one numerical variable after grouping them by one categorical variable.

Hence, I make a function to get the whole summary across one group easily, by just inputting the vector of the numerical variable and another corresponding vector of the categorical variable.

```{r}
#' Summary Statistics of Numerical Variable by Group
#'
#' Get the summary statistics (max, min, mean and median) of a numerical variable 
#' across the groups of a categorical variable.
#'
#' @param numeric_vector An vector of class "numeric", each numerical value is a part of one observation row.
#'                       To make the name of parameter be more specific, I named it as "numeric_vector".
#' @param group_vector An vector which represents the group type of each numeric value in the "numeric_vector" one by one.
#'                       To emphasize the group type, I named it as "group_vector".
#'
#' @return An object of class "data.frame" that contains the summary statistics of each group type.
#' @export
summary_by_group <- function (numeric_vector, group_vector) {
  
  # Throw an error if the first input is not numeric.
   if(!is.numeric(numeric_vector)) {
    stop('Sorry, the first argument of this function must be numeric input!\n',
         'You have provided an object of class: ', class(numeric_vector)[1])
   }
  
  # Throw an error if two inputs have different length.
   if(length(numeric_vector) != length(group_vector)){
    stop('Sorry, the length of numeric_vector and group_vector are not equal. Data frame cannot be created!\n',
         'The length of numeric_vector you provided is : ', length(numeric_vector),
         '\nThe length of group_vector you provided is : ', length(group_vector))
  }
  
  # Throw an error if the length of input is zero.
   if(length(numeric_vector) == 0){
    stop('Sorry, the length of input vector cannot be zero.')
  }
  
  # Firstly make a tibble with two inputs, drop NA values, then group them by "group_vector", compute the summary statistics.
  dplyr::tibble(numeric_vector, group_vector) %>% drop_na() %>% 
       dplyr::group_by(group_vector) %>%
       dplyr::summarise(max = max(numeric_vector), min = min(numeric_vector), 
                mean = mean(numeric_vector), median = median(numeric_vector))
}

```



## Exercise 3: Include examples (15 points)

### Examples 1

In STAT545A, I proposed a research question "Are trees with a specific genus have longer diameter?". To investigate this question, I need to get the summary statistics trees' diameters across trees' genus.

Here below shows the dataset `vancouver_trees` by selecting the `tree_id`, `genus_name` and `diameter` firstly.

```{r}
select(vancouver_trees, tree_id, genus_name, diameter, everything())
```
Now, I call the function I made in **Exercise1** above to get the the summary statistics of `diameter` across group `genus_name`.

```{r}
summary_by_group(vancouver_trees$diameter, vancouver_trees$genus_name)
```


### Examples 2

I can also create `numeric_vector` and `group_vector` by myself, as long as the values in two vectors are correspondent to each other.

```{r}
numeric_vector <- c(1.5, 7.6, 3.0, 4.7, 8, 8.4)
group_vector <- c("A", "C", "A", "B", "B", "A")
summary_by_group(numeric_vector, group_vector)
```

As we can see in the dataframe above, we get the summary statistics of three groups A, B and C.

### Examples 3

If `numeric_vector` and `group_vector` have the same length but with NA values, the function `summary_by_group` will drop NA values and then compute the summary statistics.

```{r}
numeric_vector <- c(1.5, NA, 3.0, 4.7, 8, 8.4)
group_vector <- c("A", "C", "A", "B", NA, "A")
summary_by_group(numeric_vector, group_vector)
```


### Examples 4

If `numeric_vector` and `group_vector` have different length, the function `summary_by_group` will throw an error and stop, showing the exact lengths of those two inputs.

```{r, error = TRUE}
numeric_vector <- c(1.5, 7.6, 3.0, 4.7, 8)
group_vector <- c("A", "C", "A", "B", "B", "A")
summary_by_group(numeric_vector, group_vector)
```


### Examples 5

If `numeric_vector` is not a numeric vector, the function `summary_by_group` will throw an error and stop, showing the provided input type.

```{r, error = TRUE}
wrong_numeric_vector <- c("1.5", "7.6", "3.0", "4.7", "8")
group_vector <- c("A", "C", "A", "B", "B", "A")
summary_by_group(wrong_numeric_vector, group_vector)
```





## Exercise 4: Test the Function (25 points)

### 4.1 Vectors with no NA’s

```{r}
numeric_vector <- c(98, 100, 78, 68, 90, 82)
group_vector <- c("Group1", "Group2", "Group3", "Group3", "Group1", "Group1")

test_that("Vectors with no NA’s", {
  
  #Test the result is a data frame.
  expect_equal('data.frame',
                    class(summary_by_group(numeric_vector, group_vector))[3])
  
  #Test the data frame has the expected columns.
  expect_true(all(c("group_vector", "max", "min", "mean", "median") %in% 
                     names(summary_by_group(numeric_vector, group_vector))))
  
  #Test all groups are included in the result.
  expect_true(all(c("Group1", "Group2", "Group3") %in% 
                    summary_by_group(numeric_vector, group_vector)$group_vector))
  
  #Test the calculated mean value for three groups are correct.
  expect_equal(c(90, 100, 73), 
                    summary_by_group(numeric_vector, group_vector)$mean)
})
```


### 4.2 Vectors that has NA’s

```{r}
numeric_vector <- c(98, NA, 78, 68, 90, 82)
group_vector <- c("Group1", "Group2", "Group3", NA, "Group1", "Group1")

test_that("Vectors that has NA’s", {
    
  #Test only Group1 and Group3 are included in the result.
  expect_true(all(c("Group1", "Group3") %in% 
                    summary_by_group(numeric_vector, group_vector)$group_vector))
  
  #Test the calculated mean value for two groups without NA values are correct.
  expect_equal(c(90, 78), 
                    summary_by_group(numeric_vector, group_vector)$mean)
})
```


### 4.3 Vectors of a different type

```{r}
wrong_numeric_vector <- c("98", "100", 78, 68, 90, 82)
group_vector <- c("Group1", "Group2", "Group3", "Group3", "Group1", "Group1")

test_that("Two input vectors have different length", {
  expect_error(summary_by_group(wrong_numeric_vector, group_vector), "Sorry, the first argument of this function must be numeric input!")
  
})
```



### 4.4 Vectors of wrong lengths

#### 4.4.1 Vectors of different length
```{r}
numeric_vector <- c(98)
group_vector <- c("Group1", "Group2")

test_that("Two input vectors have different length", {
  expect_error(summary_by_group(numeric_vector, group_vector), "Sorry, the length of numeric_vector and group_vector are not equal. Data frame cannot be created!")
  
})
```

#### 4.4.2 Vectors of length 0

```{r}
test_that("Two input vectors have length 0", {
  expect_error(summary_by_group(numeric(0), numeric(0)), "Sorry, the length of input vector cannot be zero.")
  
})
```



