---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# yueshen

<!-- badges: start -->
<!-- badges: end -->

The goal of package `yueshen` is to get the summary statistics (max, min, mean and median) of a numerical variable across the groups of a categorical variable.

## Installation

You can install the released version of `yueshen` from [CRAN](https://CRAN.R-project.org) with:

``` r
devtools::install_github("stat545ubc-2021/functions-SylvieShen")
```


## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(yueshen)
summary_by_group(c(1.5, 7.6, 3.0, 4.7, 8, 8.4), c("A", "C", "A", "B", "B", "A"))
```

The following example outputs the summary statistics of the numerical variable `diameter` across the groups of `genus_name` in the dataset `vancouver_trees`

```{r}
summary_by_group(datateachr::vancouver_trees$diameter, datateachr::vancouver_trees$genus_name)
```

For more information, please refer to the development documentation of function by running ```?summary_by_group``` in the **Console** of **RStudio**.
